---
title: "Intro to Geocoding in R"
author: "Yuxiao Luo"
date: "2023-02-15"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Concept

### Geocoding
What is the meaning of geocoding? Geocoding is the process of transforming a description of a location—such as a pair of coordinates, an address, or a name of a place—to a location on the earth's surface. You can geocode by entering one location description at a time or by providing many of them at once in a table.

### Reverse-geocoding
Reverse geocoding is a way to convert latitude and longitude coordinates to addresses. It’s the opposite of geocoding which lets you find latitude and longitude by address. All kinds of companies use reverse geocoding to find accurate address data for things like delivery updates, location-based advertising, or as part of an address verification solution.

We can use the`tidygeocoder` package in R to do both work. 

## Code Demo

### Geocoding
First, let's install the package and load it into R environment. 

```{r load package, message=FALSE}
# install the package if you don't have it
# install.packages('tidygeocoder')
library(tidygeocoder)
library(tidyverse)
```

Create a dataframe with addresses

```{r}
addresses <- tibble::tribble(
~name,                  ~addr,
"White House",          "1600 Pennsylvania Ave NW, Washington, DC",
"Transamerica Pyramid", "600 Montgomery St, San Francisco, CA 94111",     
"Willis Tower",         "233 S Wacker Dr, Chicago, IL 60606"                                  
)
```

This is the look of the `addresses` data frame/[tibble](https://tibble.tidyverse.org/).  

```{r}
addresses
```

Geocode the addresses using `geocode` function. The returned message from the API query contains many columns, including class of the address, type of the address, license information, and coordinates. Let's extract the latitudes and longitude in the return message with `lat = latitude` and `long = longtitude`. 

```{r, message=FALSE}
# geocode the addresses
lat_longs <- addresses %>%
  geocode(addr, method = 'osm', lat = latitude , long = longitude)

# this result only returns latitudes and longitudes
lat_longs
```

If you need more information from the address that is provided from the API serivce, you can use the argument `full_results = TRUE`. Notice that the address details may vary from different API, which is the `method = 'osm'` argument you specify in the function. 

```{r, message=FALSE}
full_result <- addresses %>% geocode(addr, method = 'osm', full_results = TRUE)
# overview of the returned full results
glimpse(full_result)
```

In the returned full results (14 columns), you can see extra address information other than  the coordinates. 

There are many geocoding services you can choose. Here is a [full list](https://jessecambon.github.io/tidygeocoder/articles/geocoder_services.html) of the services offered in `tidygeocoder`. I tested with `osm` (OpenStreetMap provided by Nominatim API), `ArcGIS` (ArcGIS) and `Bing` (Microsoft Bing Map), and here is my feedback:

- `osm` can return complete and reliable result, the speed limit is 1 query per second. It doesn't need API key. 

- `ArcGIS` can return the primary results in a fast and more efficient manner, but some information might be missing, ex., postal code. However, it doesn't have any usage limitation as far as I know. It doesn't require an API key either. 

- `Bing` can return complete and reliable result with a nice speed. It needs [registration](https://www.bingmapsportal.com/) and requires API key. It offers a free quota for trial usage (30000 queries). However, if you are working on a research project in a university or use the API queries for educational purpose, you can apply for education license, which will give you 30000 queries per day. You can find the instruction and email address in the Bing Developer Portal for applying to eudcational license after you register with the link above.

### Reverse Geocoding

To do reverse geocoding, we need geolocation coordinates (latitude and longitude). And then we can find the address and assoicated details of that address. 

I downloaded the crime reports in NYC from the [NYC Open Data webiste](https://opendata.cityofnewyork.us/), and subset 5 observations from the large dataset. I anonymized the data by replacing the original complaint ID with a randomly generated ID for each report and removing the dates. 

First, let's load the data into R environment. We can load it from the RUG GitHub folder. 

```{r, load address, message=FALSE}
address <- read_csv('https://raw.githubusercontent.com/YuxiaoLuo/RUG-RUserGroup/main/data/address.csv')
```

Let's glimpse the data. OFNS_DESC is the offense description. PD_DESC is the police department standard description of the crime. We will use the **Latitude** and **Longitude** to obtain the detailed address. 

```{r}
glimpse(address)
```

Reverse geocoding uses `reverse_geocode` function. You need to specify the latitude and longitude columns for the arguments `lat` and `long`. And specify the which API service you will use in `method` argument, here I'm using `osm`. 

```{r, message = FALSE}
address_osm <- reverse_geocode(address, 
                               lat = Latitude, 
                               long = Longitude, 
                               method = 'osm')
glimpse(address_osm)
```

The result will return the address associated with the coordinates. If you want more details with the address. You can specify `full_result = TRUE`. 

```{r reverse-full, message = FALSE}
address_osm_full <- reverse_geocode(address, Latitude, Longitude, method = 'osm', full_results = TRUE)
glimpse(address_osm_full)
```

I can extract whatever information I need in the full results. For example, I need the county, city, and postcode of the coordinates.

```{r reverse-info, message=FALSE}
address_osm_full_subset <- address_osm_full %>% select(id, county, city, postcode)
address_osm_full_subset 
```

### API key
If you have a large amount of geocoding queries, I would recommend to use any of service that requires API key since it's providing a stable and fast service. Assume you've registered any of the service and obtained an API key. You can set up the API key in `Sys.setenv()` by replacing `***Your API key***` with your own unique key (a series of digits and letters). Please keep your API key in a secret and a safe place. 

```{r API key, eval = FALSE}
Sys.setenv(HERE_API_KEY = '***Your API key***')
```
And then when you run `geocode` or `reverse_geocode` functions, choose the corresponding API service method. For example, I'm assuming you are using Google API service, then you should use `method = 'google'`.

```{r eval = FALSE}
geocode(some_addresses, address = addr, method = 'google', lat = latitude, long = longitude)
```

That's it. I hope you enjoyed reading this tutorial for Geocoding in R. 
