---
title: "Intro to Geocoding in R"
author: "Yuxiao Luo"
date: "2023-02-15"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Concept

### Geocoding
What is the meaning of geocoding? Geocoding is the process of transforming a description of a location—such as a pair of coordinates, an address, or a name of a place—to a location on the earth's surface. You can geocode by entering one location description at a time or by providing many of them at once in a table.

### Reverse-geocoding
Reverse geocoding is a way to convert latitude and longitude coordinates to addresses. It’s the opposite of geocoding which lets you find latitude and longitude by address. All kinds of companies use reverse geocoding to find accurate address data for things like delivery updates, location-based advertising, or as part of an address verification solution.

We can use the`tidygeocoder` package in R to do both work. 

## Code Demo

### Geocoding
First, let's install the package and load it into R environment. 

```{r load package, message=FALSE}
# install the package if you don't have it
# install.packages('tidygeocoder')
library(tidygeocoder)
library(tidyverse)
```

Create a dataframe with addresses

```{r}
addresses <- tibble::tribble(
~name,                  ~addr,
"White House",          "1600 Pennsylvania Ave NW, Washington, DC",
"Transamerica Pyramid", "600 Montgomery St, San Francisco, CA 94111",     
"Willis Tower",         "233 S Wacker Dr, Chicago, IL 60606"                                  
)
addresses
```

Geocode the addresses

```{r, message=FALSE}
# geocode the addresses
lat_longs <- addresses %>%
  geocode(addr, method = 'osm', lat = latitude , long = longitude)
lat_longs
```

### Reverse Geocoding

To do reverse geocoding, we need geolocation coordinates (latitude and longitude). And then we can find the address and assoicated details of that address. 

I downloaded the crime reports in NYC from the [NYC Open Data webiste](https://opendata.cityofnewyork.us/), and subset 5 observations from the large dataset. I anonymized the data by replacing the original complaint ID with a randomly generated ID for each report and removing the dates. 

First, let's load the data into R environment and take a look at it.

```{r, load address}

```




